---
title: "Cross-Validation"
output: html_notebook
---

This is a script uses different cross validation methods to interpret the predictive
of our Python Script.

### Set directory
```{r}
setwd("C:\\ISE\\Senior_Design\\Cross Validation")
```

### Load packages

```{r}
library(tidyverse) # package to perform data manipulation and visualization
library(caret) #  package to compute cross - validation methods
library(readxl) # package to read excel files
library(dplyr) # package to compute random forest method
library(ggplot2) # package to plot bar chart
library(MLmetrics) # package for MAPE
```

### Load data

```{r}
cow_data <- read_excel("Cross_Validation_Data_v4.xlsx",sheet="All factors")
```

### Check dataset

```{r}
head(cow_data)
dim(cow_data)
```

### Setting seed to generate a reproducible random sampling

```{r}
set.seed(123)
```

### Creating training data using 80%, 90%, and 70% of the dataset

```{r}
random_sample_80 <- createDataPartition(cow_data $ Given,p = 0.8, list = FALSE)
random_sample_90 <- createDataPartition(cow_data $ Given,p = 0.9, list = FALSE)
random_sample_70 <- createDataPartition(cow_data $ Given,p = 0.7, list = FALSE)
```

### Generating training dataset from the random_sample

```{r}
training_dataset_80  <- cow_data[random_sample_80, ]
training_dataset_90  <- cow_data[random_sample_90, ]
training_dataset_70  <- cow_data[random_sample_70, ]
dim(training_dataset_80)
```

### Generating testing dataset from rows which are not included in random_sample

```{r}
testing_dataset_80 <- cow_data[-random_sample_80, ]
testing_dataset_90 <- cow_data[-random_sample_90, ]
testing_dataset_70 <- cow_data[-random_sample_70, ]
dim(testing_dataset_80)
```

## Building the model

### Training the model by assigning Given column as target variable and rest other columns as independent variables

```{r}
model_80 <- lm(Given ~., data = training_dataset_80)
model_90 <- lm(Given ~., data = training_dataset_90)
model_70 <- lm(Given ~., data = training_dataset_70)

```

### Predicting the target variable

```{r}
predictions_80 <- predict(model_80, testing_dataset_80)
predictions_90 <- predict(model_90, testing_dataset_90)
predictions_70 <- predict(model_70, testing_dataset_70)

```

### Computing model performance metrics

```{r}
data.frame( R2 = R2(predictions_80, testing_dataset_80 $ Given),
            RMSE = RMSE(predictions_80, testing_dataset_80 $ Given),
            MAE = MAE(predictions_80, testing_dataset_80 $ Given),
            MAPE = MAPE(predictions_80,testing_dataset_80 $ Given))
data.frame( R2 = R2(predictions_90, testing_dataset_90 $ Given),
            RMSE = RMSE(predictions_90, testing_dataset_90 $ Given),
            MAE = MAE(predictions_90, testing_dataset_90 $ Given),
            MAPE = MAPE(predictions_90, testing_dataset_90 $ Given))
data.frame( R2 = R2(predictions_70, testing_dataset_70 $ Given),
            RMSE = RMSE(predictions_70, testing_dataset_70 $ Given),
            MAE = MAE(predictions_70, testing_dataset_70 $ Given),
            MAPE = MAPE(predictions_70, testing_dataset_70 $ Given))
```

## Leave one out cross validation
 
### Defining training control as Leave One Out Cross Validation

```{r}
train_control <- trainControl(method = "LOOCV")
```

### Training the model by assigning Given column as target variable and rest other column as independent variable

```{r}
LOOCV_model <- train(Given ~., data = cow_data,
               method = "lm",
               trControl = train_control)
```

### Print model performance metrics along with other details

```{r}
print(LOOCV_model)
```

## K-fold cross-validation
 
### Setting seed to generate a reproducible random sampling
```{r}
set.seed(125)
```

### Defining training control as cross-validation and value of K equal to 10

```{r}
train_control <- trainControl(method = "cv",number = 10)
```

### Training the model by assigning Given column as target variable and rest other column as independent variable

```{r}
K_model <- train(Given ~., data = cow_data,
               method = "lm",
               trControl = train_control)
```

### Printing model performance metrics along with other details

```{r}
print(K_model)
```

## Repeated K-fold cross-validation

### Setting seed to generate a reproducible random sampling
```{r}
set.seed(125)
```

### Defining training control as repeated cross-validation and value of K is 10 and repetition is 3 times

```{r}
train_control <- trainControl(method = "repeatedcv",
                            number = 10, repeats = 3)
```

### Training the model by assigning sales column as target variable and rest other column as independent variable

```{r}
Repeat_k_model <- train(Given ~., data = cow_data,
               method = "lm",
               trControl = train_control)
```

### Printing model performance metrics along with other details
```{r}
print(Repeat_k_model)
```

## Random Forest method

### Setting seed to generate a reproducible random sampling 
```{r}
set.seed(123)
```

### Defining random forest model

```{r}
forest <- train(Given~., data=training_dataset_80, method='rf', 
        trControl=train_control)
```

### Print out the details about the model

```{r}
forest$finalModel
```

### Get variable importance, and turn into a data frame

```{r}
var_imp <- varImp(forest, scale=FALSE)$importance
var_imp <- data.frame(variables=row.names(var_imp), importance=var_imp$Overall)
```

### Create a plot of variable importance
```{r}
var_imp %>%
    arrange(importance) %>%
    ggplot(aes(x=reorder(variables, importance), y=importance)) + 
    geom_bar(stat='identity') + 
    coord_flip() + 
    xlab('Variables') +
    labs(title='Random forest variable importance') + 
    theme_minimal() + 
    theme(axis.text = element_text(size = 10), 
        axis.title = element_text(size = 15), 
        plot.title = element_text(size = 20))
```

### Generate predictions

```{r}
forest_predictions <- predict(object=forest, newdata=testing_dataset_80[, -1])
data.frame( R2 = R2(forest_predictions, testing_dataset_80 $ Given),
            RMSE = RMSE(forest_predictions, testing_dataset_80 $ Given),
            MAE = MAE(forest_predictions, testing_dataset_80 $ Given),
            MAPE = MAPE(forest_predictions, testing_dataset_80 $ Given))
```
